---
title: "P8105 Homework 3 Solutions"
author: Meghan Bellerose
date: October 1, 2020
output: html_document
---

```{r, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1: Instacart

First, I'll load Instracart data from the p8105 library.

```{r}
library(p8105.datasets)
data("instacart")
```

The Instracart dataset...

There are X aisles. The aisles that most items are ordered from are X, X, and X.


The following plot shows the number of items ordered in each aisle (limited to aisles with more than 1000 items ordered). 

This table shows the three most popular items in the aisles: “baking ingredients”, “dog food care”, and “packaged vegetables fruits” and how many times each item is ordered. 

This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

# Problem 2: Accelerometers 



# Problem 3

